---
title: 'Assignment-2 :: Fundamentals of Computational Mathematics'
author: "Chirag Vithalani"
date: "September 4, 2016"
output: html_document
---

<hr />

#####1 Show that A^T A 6 $\neq$ AA^T in general.

Let's say <br /><br /><br />
<math xmlns="http://www.w3.org/1998/Math/MathML"><mi>A</mi><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>a</mi></mtd></mtr><mtr><mtd><mi>b</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mspace linebreak="newline"/><mi>A</mi><msup><mi>A</mi><mi>T</mi></msup><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>a</mi></mtd></mtr><mtr><mtd><mi>b</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mo>&#xA0;</mo><mo>&#xA0;</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>b</mi></mtd></mtr><mtr><mtd><mi>a</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mo>&#xA0;</mo><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn><mo>+</mo><msup><mi>a</mi><mn>2</mn></msup></mtd><mtd><mi>b</mi><mo>+</mo><mi>a</mi></mtd></mtr><mtr><mtd><mi>b</mi><mo>+</mo><mi>a</mi></mtd><mtd><msup><mi>b</mi><mn>2</mn></msup><mo>+</mo><mn>1</mn></mtd></mtr></mtable></mfenced><mspace linebreak="newline"/><msup><mi>A</mi><mi>T</mi></msup><mi>A</mi><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>b</mi></mtd></mtr><mtr><mtd><mi>a</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mo>&#xA0;</mo><mo>&#xA0;</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>a</mi></mtd></mtr><mtr><mtd><mi>b</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mo>&#xA0;</mo><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn><mo>+</mo><msup><mi>b</mi><mn>2</mn></msup></mtd><mtd><mi>a</mi><mo>+</mo><mi>b</mi></mtd></mtr><mtr><mtd><mi>a</mi><mo>+</mo><mi>b</mi></mtd><mtd><msup><mi>a</mi><mn>2</mn></msup><mo>+</mo><mn>1</mn></mtd></mtr></mtable></mfenced><mspace linebreak="newline"/><mi>A</mi><msup><mi>A</mi><mi>T</mi></msup><mo>&#x2260;</mo><msup><mi>A</mi><mi>T</mi></msup><mi>A</mi><mspace linebreak="newline"/></math>

<b>Using R</b>

```{r}
A=matrix(c(1,2,3,4), nrow=2, ncol=2) 
A
t(A)
A %*% t(A)
t(A) %*% A

```

#####2. For a special type of square matrix A, we get AT A = AAT.Under what conditions could this be true?

For the identity matrix and  its variants, and symetric matrices $A^TA = AA^T$. Because a symmetric matrix and its transpose are equal, that is multiplying a matrix by itself.

#####3. Write an R function to factorize a square matrix $A$ into LU or LDU 
This function will factorize matrix into LU
```{r}
factorize <- function(A) 
{
  if (nrow(A) == ncol(A)) 
  {
      size <- nrow(A)
      L <- diag(size) #construct a diagonal matrix.
      
      
      for (i in 1:(size - 1)) 
      {
          for (j in (i + 1):size) 
          {
              L[j, i] <- A[j, i] / A[i, i]
              A[j, ]  <- A[j, ] - L[j, i] * A[i, ]
          }
      }
      print("check if below is same as original matrix")
      print("******L %*% A********")
      print(L %*% A)
      print("******L********")
      print(L)
      print("******U********")
      print(A)
  } 
}
```

Check with a 2x2 and a 3x3 matrix
```{r}
# 2x2
A2x2 <- matrix(c(2,3,4,5), nrow = 2)
A2x2
factorize(A2x2)

# 3x3
A3x3 <- matrix(c(8,9,10,0,1,2,11,12,13), nrow = 3)
A3x3
factorize(A3x3)
```


