---
title: 'Assignment-2 :: Fundamentals of Computational Mathematics'
author: "Chirag Vithalani"
date: "September 4, 2016"
output: html_document
---

<hr />

#####1 Show that A^T A 6 $\neq$ AA^T in general.

Let's say <br /><br /><br />
<math xmlns="http://www.w3.org/1998/Math/MathML"><mi>A</mi><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>a</mi></mtd></mtr><mtr><mtd><mi>b</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mspace linebreak="newline"/><mi>A</mi><msup><mi>A</mi><mi>T</mi></msup><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>a</mi></mtd></mtr><mtr><mtd><mi>b</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mo>&#xA0;</mo><mo>&#xA0;</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>b</mi></mtd></mtr><mtr><mtd><mi>a</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mo>&#xA0;</mo><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn><mo>+</mo><msup><mi>a</mi><mn>2</mn></msup></mtd><mtd><mi>b</mi><mo>+</mo><mi>a</mi></mtd></mtr><mtr><mtd><mi>b</mi><mo>+</mo><mi>a</mi></mtd><mtd><msup><mi>b</mi><mn>2</mn></msup><mo>+</mo><mn>1</mn></mtd></mtr></mtable></mfenced><mspace linebreak="newline"/><msup><mi>A</mi><mi>T</mi></msup><mi>A</mi><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>b</mi></mtd></mtr><mtr><mtd><mi>a</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mo>&#xA0;</mo><mo>&#xA0;</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn></mtd><mtd><mi>a</mi></mtd></mtr><mtr><mtd><mi>b</mi></mtd><mtd><mn>1</mn></mtd></mtr></mtable></mfenced><mo>&#xA0;</mo><mo>=</mo><mfenced open="|" close="|"><mtable><mtr><mtd><mn>1</mn><mo>+</mo><msup><mi>b</mi><mn>2</mn></msup></mtd><mtd><mi>a</mi><mo>+</mo><mi>b</mi></mtd></mtr><mtr><mtd><mi>a</mi><mo>+</mo><mi>b</mi></mtd><mtd><msup><mi>a</mi><mn>2</mn></msup><mo>+</mo><mn>1</mn></mtd></mtr></mtable></mfenced><mspace linebreak="newline"/><mi>A</mi><msup><mi>A</mi><mi>T</mi></msup><mo>&#x2260;</mo><msup><mi>A</mi><mi>T</mi></msup><mi>A</mi><mspace linebreak="newline"/></math>

<b>Using R</b>

```{r}
A=matrix(c(1,2,3,4), nrow=2, ncol=2) 
A
t(A)
A %*% t(A)
t(A) %*% A

```

#####2. For a special type of square matrix A, we get AT A = AAT.Under what conditions could this be true?

For the identity matrix and  its variants, and symetric matrices $A^TA = AA^T$. Because a symmetric matrix and its transpose are equal, that is multiplying a matrix by itself.

#####3. Write an R function to factorize a square matrix $A$ into LU or LDU 
This function will factorize a 2x2 and a 3x3 matrix into LU
```{r}
f_LU <- function(A) {

#Check dimensions of the matrix - 2x2 or 3x3

# 2x2
if (nrow(A) == 2) {

  #Subract multipler
  E <- -A[2,1] / A[1,1]
  L <- matrix(c(1,0,E,1),nrow=2)
  U <- L %*% A

  L <- solve(L)
  
  print ("------ A ------")
  print (A)
  print ("------ L ------")
  print(L)
  print ("------ U ------")
  print(U)

  (L %*% U == A)

#3x3
} else if (nrow(A) == 3) {

  #Subract multipler
  e_21 <- -A[2,1] / A[1,1]
  me_21 <- matrix(c(1,0,0,e_21,1,0,0,0,1), nrow=3)
  A2 <- me_21 %*% A

  #Subract multipler
  e_31 <- -A2[3,1] / A2[1,1]
  me_31 <- matrix(c(1,0,0,0,1,0,e_31,0,1), nrow=3)
  A3 <- me_31 %*% A2

  #Subtract multipler
  e_32 <- -A3[3,2] / A3[2,2]
  me_32 <- matrix(c(1,0,0,0,1,0,0,e_32,1), nrow=3)
  U <- me_32 %*% A3

  L <- solve(me_21) %*% solve(me_31) %*% solve(me_32)

  A <- L %*% U
  
  print ("------ A ------")
  print(A)
  print ("------ L ------")
  print(L)
  print ("------ U ------")
  print(U)

  (L %*% U == A)
 
  }
}
```

Check with a 2x2 and a 3x3 matrix
```{r}
# 2x2
A2x2 <- matrix(c(2,3,4,5), nrow = 2)
f_LU(A2x2)

# 3x3
A3x3 <- matrix(c(8,9,10,0,1,2,11,12,13), nrow = 3)
f_LU(A3x3)
```


