---
title: 'Assignment-1 :: Fundamentals of Computational Mathematics'
author: "Chirag Vithalani"
date: "August 31, 2016"
output: html_document
---
<hr />

####Problem Set 1:
#####(1) Calculate the dot product of $u.v$ where $u=[0.5;0.5]$ and $v=[3;-4]$

```{r}

u <-  c(0.5,0.5)
v <- c(3,-4)

dot.product<-u%*%v
dot.product
```

#####(2) What are the lengths of $u$ and $v$? 

Length of u:
```{r}
Length_of_u<-sqrt(u %*% u)
Length_of_u
```
Length of v:
```{r}
Length_of_v<-sqrt(v %*% v)
Length_of_v
```

#####(3) What is the linear combination: $3u-2v$?

```{r}
3*u -2*v

```
#####(4) What is the angle between u and v

```{r}
theta <- acos(dot.product/(Length_of_u*Length_of_v))
theta
```

####Problem Set 2:
Set up a system of equations with 3 variables and 3 constraints and solve for x. Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination. Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system. You don't have to worry about degenerate cases and can safely assume that the function will only be tested with a system of equations that has a solution. Please note that you do have to worry about zero pivots, though. Please note that you should not use the built-in function solve to solve this system or use matrix inverses. The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure.

Please test it with the system below and it should produce a solution $x = [-1.55;-0.32;0.95]$
$$\begin{bmatrix}1 & 1 & 3 \\2 & -1 & 5 \\-1 & -2 & 4\end{bmatrix}\,\begin{bmatrix}x_1 \\x_2 \\x_3\end{bmatrix}=\begin{bmatrix}1 \\2 \\6\end{bmatrix}$$  



```{r}
# from http://stackoverflow.com/questions/16044377/how-to-do-gaussian-elimination-in-r-do-not-use-solve

A <- array(c(1, 2, -1, 1, -1, -2, 3, 5, 4), dim=c(3,3))
b <- c(1, 2, 6) # constraint vector

solve_this<-function (A,b)
{
  p <- nrow(A)
  
  # binding A and b
  binded <- cbind(A,b)  
  binded[1,] <- binded[1,]/binded[1,1]
  
  for (i in 2:p)
  {
	   for (j in i:p) 
	   {
			binded[j, ] <- binded[j, ] - binded[i-1, ] * binded[j, i-1]
	   }
	   binded[i,] <- binded[i,]/binded[i,i]
  }
  
  for (i in p:2)
  {
	   for (j in i:2-1) 
	   {
			binded[j, ] <- binded[j, ] - binded[i, ] * binded[j, i]
	   }
  }
  print (binded)
}

solve_this(A,b)

```

####### Verifying using `solve` function:
```{r}
A <- array(c(1, 2, -1, 1, -1, -2, 3, 5, 4), dim=c(3,3))
b <- c(1, 2, 6)
solve(A, b)


